D.I.A.G.R.A.M.: Development of Image Analysis for Graph Recognition And Modeling

### 2025-05-26 doc/article/article.tex
Device

@@

\section{Experiments}
\subsection{Text Digitization}
In order to choose an appropriate model, some tests were done over our specific dataset employing some different, pre-made models. The average results over the dataset are shown in Table~\ref{tab:text_digitization}.

\begin{table}[htbp]
\caption{Text Distance Metrics}
\centering
\begin{tabular}{lccc}
\hline
\textbf{Model} & \textbf{Hamming} & \textbf{Cosine} & \textbf{Euclidean} \\
\hline
microsoft-trocr-small-printed & 1.392 & 0.150 & 7133 \\
% microsoft-trocr-base-printed & 1.286 & 0.081 & 3594 \\
microsoft-trocr-small-handwritten & \textbf{1.346} & 0.100  & 3947 \\
microsoft-trocr-base-handwritten & 1.549 & \textbf{0.057} & \textbf{3003} \\
\hline
\end{tabular}
\label{tab:text_digitization}
\end{table}

>>>>

\subsubsection{Arrow Orientation}
To later correctly assign arrows to their source and target elements, the need to determine where the tip of an arrow lies. This task involves a few problems, among which determining exactly the arrow shape and its end - because sometimes other pen signs may be found in the bounding box of the arrow itself. [TODO: Dire l'approccio finale usato]. \\

@@

\section{Experiments}

\subsection{Classifier}
TODO

\subsection{Object Detection}
TODO

\subsection{Arrow Orientation}
As we've seen in the previous section, the need to find the exact position of an arrow's head arises.
The first algorithmic approach proposed to solve this task - called \textit{Double Clustering} - was as follows: 
\begin{enumerate}
	\item Apply a clustering algorithm to the bounding box.
	\item Suppose that the biggest cluster belongs to the arrow; the other clusters are thought to contain useless elements for this task, and can be discarded.
	\item Find the keypoints of the arrow using SIFT.
	\item Find the two keypoints inside the arrow with maximum distance. Assume that these two points are closely related to the two ends of the arrow.
	\item Run a template matching procedure inside the two clusters to find the head of the arrow.
\end{enumerate}

For Step (1), we tried to employ both Spectral Clustering (see Figure~\ref{fig:spectral_clustering}) and DBSCan (see Figure~\ref{fig:arrow_double_clustering_clusters}).

\begin{figure}[H]
\centering
\includegraphics[width=150pt]{spectral_clustering.png}
\caption{The result of Spectral Clustering to one instance of an arrow bounding box in our dataset.}
\label{fig:spectral_clustering}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=150pt]{arrow_double_clustering_clusters.png}
\caption{The result of DBScan to one instance of an arrow bounding box in our dataset.}
\label{fig:arrow_double_clustering_clusters}
\end{figure}

For Step (3), we used SIFT in order to find keypoints and some descriptors for them - necessary for the later template matching.
During Step (4), we were able to correctly deduce the overall direction of the arrow (see Figure~\ref{fig:arrow_double_clustering_direction}).

\begin{figure}[H]
\centering
\includegraphics[width=150pt]{arrow_double_clustering_direction.png}
\caption{The estimated direction of the arrow, finding the points that maximize the distance inside the cluster.}
\label{fig:arrow_double_clustering_direction}
\end{figure}

For Step (5), we had previously hand-picked some templates (see Figure~\ref{fig:templates}).

\begin{figure}[H]
\centering
\includegraphics[width=50pt]{templates.png}
\caption{An example of hand-picked template with SIFT descriptors.}
\label{fig:templates}
\end{figure}

However, this approach didn't work as expected (see Figure~\ref{fig:template_matching_fail}). [TODO: Spiegare perchè].

\begin{figure}[H]
\centering
\includegraphics[width=150pt]{template_matching_fail.png}
\caption{The overall result of the algorithm.}
\label{fig:template_matching_fail}
\end{figure}

\subsection{Text Digitization}
In order to choose an appropriate model, some tests were done over our specific dataset employing some different, pre-made models. The average results over the dataset are shown in Table~\ref{tab:text_digitization}.

\begin{table}[htbp]
\caption{Text Distance Metrics}
\centering
\begin{tabular}{lccc}
\hline
\textbf{Model} & \textbf{Hamming} & \textbf{Cosine} & \textbf{Euclidean} \\
\hline
microsoft-trocr-small-printed & 1.392 & 0.150 & 7133 \\
% microsoft-trocr-base-printed & 1.286 & 0.081 & 3594 \\
microsoft-trocr-small-handwritten & \textbf{1.346} & 0.100  & 3947 \\ \hline \textbf{microsoft-trocr-base-handwritten} & 1.549 & \textbf{0.057} & \textbf{3003} \\
\hline
\end{tabular}
\label{tab:text_digitization}
\end{table}
